\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-05-10 forwards-and-backwards.tex
ian and terence comments

	\comment{Need to thank KAVLI, Paris Cit\'e, various funders}

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction-diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation 
(that is equation~(\ref{semilinear PDE}) with $\DG=\Delta$ and $F(\varphi)=1-\varphi$)
emerges as a special case.

@@

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f \in C_b^\infty(\IR^d)$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align*}
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    % \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align*}
\end{definition}

@@

{\bf Dispersal generator:}
We assume that
\begin{enumerate}
    \item $\meanq(x)$ and $\covq(x)$ are
        $\alpha$-H\"older continuous for some $\alpha \in (0, 1]$
        and uniformly bounded
        in each component;
    % \item $\meanq(x)$ and $\covq(x)$ are $\alpha$-H\"older continuous in each component, for some $\alpha \in (0,1]$;
    \item the operator $\DG$ is uniformly strictly elliptic.
\end{enumerate}

@@

Specialising the results of 
\citet{kurtz/xiong:1999} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \citet{kurtz/xiong:1999} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
in a potential tilted by migration bias
($(\varphi_s \gamma)^2 \exp(-\meanq \cdot x / \sigma^2)$).
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/fig2a}
        \includegraphics[width=0.7\textwidth]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\paragraph{Porous Medium equation with logistic growth:}
Finally, consider equation~(\ref{PME}). Setting $x_0=0$ (for definiteness) and substituting
the form of $w^P$ from equation~\eqref{eqn:pme_wave}
into Corollary~\ref{cor:wavefront},
with $c=1$,
$\gamma(x, m) = m$,
$r(x,w) = 1$,
and $F(x, m) = (1 - m)$,
the generator of the diffusion governing the position of the ancestral lineage relative to the wavefront
is, for $x < 0$,
%with $f_x$ denoting the spatial derivative of $f$,
\begin{align*}
    \Lgen_P f
    &=
        w^P
        \left(
        \partial_{xx} f
         +
         2 \frac{\partial_x((w^P)^2)}{(w^P)^2} \partial_xf
        \right)
        + \partial_xf \\
    &=
        \left(1 - e^{\frac{1}{2} x} \right)
        \partial_{xx}f
        -
        2 e^{\frac{1}{2} x} \partial_xf
        +
        \partial_xf .
\end{align*}
The speed measure corresponding to this diffusion has density
%again for $\xi < 0$,
\begin{align*}
    m_P(\xi)
    &\propto
        \frac{ 1 }{ 2 (1 - e^{\xi/2}) }
        \exp\left(
            \int_\eta^\xi \left\{
                1 - \frac{e^{x/2}}{1 - e^{x/2}}
            \right\} dx
        \right) \\
    &\propto
        e^\xi\left(1-e^{\xi/2}\right),
        \quad
        \text{for } \xi < 0 ,
\end{align*}
which is integrable and so when suitably normalised gives the unique stationary distribution.
Notice that even though we have the same reaction term as in the Fisher-KPP equation, with this 
form of nonlinear diffusion, at stationarity
the lineage will typically be significantly behind the front, 
suggesting a different genealogy.

@@

We use the fact that $\int q_\theta(x,dz)=1$ to rewrite~(\ref{mean measure})
as 
\begin{equation}
\label{eqn:rewritten mean measure}
\begin{split}
    \int\left(
        \int \theta \left( f(z) r(z,\eta)- f(x) r(x,\eta) \right) q_\theta(x,dz)
    \right)
    \gamma(x,\eta)
    \eta(dx)
    \\
    + \int f(x) \theta \Big(
        r(x,\eta) \gamma(x,\eta)
        - \mu_\theta(x,\eta)
    \Big) \eta(dx).
\end{split}
\end{equation}
We have defined $\mu_\theta$ so that the second term is simple:
\begin{align*}
    \theta \Big( r(x,\eta) \gamma(x,\eta) - \mu_\theta(x,\eta) \Big)
    = F(x, \eta) .
\end{align*}
Furthermore, recall from Definition~\ref{def:dispersal_generator} that
\begin{align} \label{eqn:near_critical}
    \int \theta \Big(
        r(z,\eta) f(z)
        -
        r(x,\eta) f(x)
    \Big) q_\theta(x,dz) 
    \qquad \stackrel{\theta\to\infty}{\longrightarrow} \qquad  
    \DG \big(r(\cdot,\eta)f(\cdot)\big)(x) .
\end{align}
In particular,
if dispersal is determined by a standard multivariate Gaussian
with mean zero and covariance $\sigma^2 I / \theta$,
then $\DG = \sigma^2 \Delta$, where $\Delta$ denotes the Laplacian.

@@

Although our proof of Theorem~\ref{thm:lineages}
uses an explicit representation in terms of the lookdown process,
the result can be understood through informal calculations.
Suppose that we have traced a lineage back to an individual 
at location $y$ at time $t$.
Looking further back through time, at the time of the birth of that 
individual, the lineage will jump to the location of the parent of the
individual.
Now, the rate at which new individuals are born to parents at $x$ 
and establish at $y$ is
$$
\theta N\eta^N_t(dx)    \gamma(x, \eta^N_t) q_\theta(x, dy) r(y, \eta^N_t) .
$$
Suppose that $\eta^N$ did have a density (in the prelimit it does not),
say $\eta^N_t(dx) = \varphi^N_t(x) dx$.
Informally,
since the number of individuals near $y$ is $N\varphi^N_t(y) dy$,
the probability that a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is
\begin{equation} \label{eqn:informal_rates}
\frac{\theta
    \varphi^N_t(x) \gamma(x, \eta^N_t) r(y, \eta^N_t)
}{
    \varphi^N_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated 
as a randomly chosen individual, we define
a continuous-time jump process 
whose transition rates, conditional on $(\varphi^N_t)_{t=0}^T$, 
are given by \eqref{eqn:informal_rates}. Because we are tracing the
lineage backwards 
in time we make the substitution $s=T-t$ and write
$(L^N_s)_{s=0}^T$ for the location of a lineage that moves 
according to these jump rates.
Then, abusing notation to write 
$q_\theta(x, y)$ for the density of $q_\theta(x,dy)$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^N_{s+ds}) - f(y) \;|\; L^N_s = y]
    \\&\qquad 
    =
    ds\, \theta \int \left(f(x) - f(y)\right)
    \frac{
        \varphi^N_{T-s}(x) \gamma(x, \eta^N_{T-s}) r(y, \eta^N_{T-s})
    }{
        \varphi^N_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
(Note that this integral is with respect to $x$.)
Referring back to Definition~\ref{def:dispersal_generator},
a quick calculation shows that as $N \to \infty$,
\begin{align*}
    &
    \theta \int \big(f(x) - f(y)\big) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \big\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \big\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

\begin{proposition}[Line of descent construction]
    \label{prop:limiting_construction}
Let 
$J(x,\eta)$ be such that $r(x,\eta)\gamma(x,\eta)\covq(x) = J(x,\eta) J(x,\eta)^T$, and set
$\beta(x, \eta) = r(x,\eta)\gamma(x,\eta)\big(\meanq(x) + 2 \covq(x) \grad \log r(x, \eta)\big).$
Associate with each label $a \in \labelspace = \cup_{k\geq 1}\IN^k$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
and with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
on $\IR^d \times [0, \infty)$ 
with mean measure $\eta_0\times \lambda$
(the product of $\eta_0$ and Lebesgue measure).
For each $i$, begin a line of descent
with label $i$, location $X_i(0) = x_i$, 
level $U_i(0) = u_i$, and birth time $\tau_i = 0$.

@@

@@

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right) .
\end{equation*}
and 
we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
The example we have in mind is $\zeta$ (and therefore $\rho$) being the density of
a mean zero Gaussian random variable.  
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}

>>>>

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction-diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation 
(that is equation~(\ref{semilinear PDE}) with $\DG=\Delta$ and $F(\varphi)=1-\varphi$)
emerges as a special case.

@@

\section*{Acknowledgements}

\comment{Need to thank KAVLI, Paris Cit\'e, various funders}
IL is supported by the ANID/Doctorado en el extranjero doctoral scholarship, grant number 2018-72190055.
TT supported by the EPSRC Centre for Doctoral Training in Mathematics of Random Systems: Analysis, Modelling and Simulation (EP/S023925/1) and the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy, EXC-2047/1-390685813.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    Furthermore, we define for $f \in C_b^\infty(\IR^d)$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    and denote the adjoint of $\DG$ by
    \begin{align*}
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    % \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align*}
\end{definition}

\begin{remark}
    \label{rem:DG_limit}
    $\DG$ is defined so that
    \begin{align*}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align*}
\end{remark}

@@

{\bf Dispersal generator:}
We assume that
\begin{enumerate}
    \item $\meanq(x)$ and $\covq(x)$ are
        $\alpha$-H\"older continuous for some $\alpha \in (0, 1]$
        and uniformly bounded
        in each component, and
    % \item $\meanq(x)$ and $\covq(x)$ are $\alpha$-H\"older continuous in each component, for some $\alpha \in (0,1]$;
    \item the operator $\DG$ is uniformly strictly elliptic,
        i.e., $\inf_x \inf_{y:\|y\|=1} \sum_{ij} y_i C(x)_{ij} y_j > 0$.
\end{enumerate}

@@

Specialising the results of 
\citet{kurtz/xiong:1999} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \citet{kurtz/xiong:1999} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \citet{kurtz/xiong:1999}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\citet{kurtz/xiong:1999} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
in a potential tilted by migration bias
($(\varphi_s \gamma)^2 \exp(-\meanq \cdot x / \sigma^2)$,
whose gradient appears in the drift term of the generator).
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/fig2a}
        \includegraphics[width=0.7\textwidth]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ in top and bottom figures
        are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\paragraph{Porous Medium equation with logistic growth:}
Finally, consider equation~(\ref{PME}). Setting $x_0=0$ (for definiteness) and substituting
the form of $w^P$ from equation~\eqref{eqn:pme_wave}
into Corollary~\ref{cor:wavefront},
with $c=1$,
$\gamma(x, m) = m$,
$r(x,w) = 1$,
and $F(x, m) = (1 - m)$,
the generator of the diffusion governing the position of the ancestral lineage relative to the wavefront
is, for $x < 0$,
%with $f_x$ denoting the spatial derivative of $f$,
\begin{align*}
    \Lgen_P f
    &=
        w^P
        \left(
        \partial_{xx} f
         +
         2 \frac{\partial_x((w^P)^2)}{(w^P)^2} \partial_xf
        \right)
        + \partial_xf \\
    &=
        \left(1 - e^{\frac{1}{2} x} \right)
        \partial_{xx}f
        -
        2 e^{\frac{1}{2} x} \partial_xf
        +
        \partial_xf .
\end{align*}
The speed measure corresponding to this diffusion has density
%again for $\xi < 0$,
\begin{align*}
    m_P(\xi)
    &\propto
        \frac{ 1 }{ 2 (1 - e^{\xi/2}) }
        \exp\left(
            \int_\eta^\xi \left\{
                1 - \frac{e^{x/2}}{1 - e^{x/2}}
            \right\} dx
        \right) \\
    &\propto
        e^\xi\left(1-e^{\xi/2}\right),
        \quad
        \text{for } \xi < 0 
\end{align*}
and $m_P(\xi) = 0$ for $\xi \ge 0$,
which is integrable and so when suitably normalised gives the unique stationary distribution.
Notice that even though we have the same reaction term as in the Fisher-KPP equation, with this 
form of nonlinear diffusion, at stationarity
the lineage will typically be significantly behind the front, 
suggesting a different genealogy.

@@

We use the fact that $\int q_\theta(x,dz)=1$ to rewrite~(\ref{mean measure})
as 
\begin{equation}
\label{eqn:rewritten mean measure}
\begin{split}
    \int\left(
        \int \theta \left( f(z) r(z,\eta)- f(x) r(x,\eta) \right) q_\theta(x,dz)
    \right)
    \gamma(x,\eta)
    \eta(dx)
    \\
    + \int f(x) \theta \Big(
        r(x,\eta) \gamma(x,\eta)
        - \mu_\theta(x,\eta)
    \Big) \eta(dx).
\end{split}
\end{equation}
We have defined $\mu_\theta$ so that the second term is simple:
\begin{align*}
    \theta \Big( r(x,\eta) \gamma(x,\eta) - \mu_\theta(x,\eta) \Big)
    = F(x, \eta) .
\end{align*}
Furthermore, recall from Remark~\ref{rem:DG_limit} that
\begin{align} \label{eqn:near_critical}
    \int \theta \Big(
        r(z,\eta) f(z)
        -
        r(x,\eta) f(x)
    \Big) q_\theta(x,dz) 
    \qquad \stackrel{\theta\to\infty}{\longrightarrow} \qquad  
    \DG \big(r(\cdot,\eta)f(\cdot)\big)(x) .
\end{align}
In particular,
if dispersal is determined by a standard multivariate Gaussian
with mean zero and covariance $\sigma^2 I / \theta$,
then $\DG = \sigma^2 \Delta$, where $\Delta$ denotes the Laplacian.

@@

Although our proof of Theorem~\ref{thm:lineages}
uses an explicit representation in terms of the lookdown process,
the result can be understood through informal calculations.
Suppose that we have traced a lineage back to an individual 
at location $y$ at time $t$.
Looking further back through time, at the time of the birth of that 
individual, the lineage will jump to the location of the parent of the
individual.
Now, the rate at which new individuals are born to parents at $x$ 
and establish at $y$ is
$$
\theta N\eta^N_t(dx)    \gamma(x, \eta^N_t) q_\theta(x, dy) r(y, \eta^N_t) .
$$
Suppose that $\eta^N$ did have a density (in the prelimit it does not),
say $\eta^N_t(dx) = \varphi^N_t(x) dx$.
Informally,
since the number of individuals near $y$ is $N\varphi^N_t(y) dy$,
the probability that a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is
\begin{equation} \label{eqn:informal_rates}
\frac{\theta
    \varphi^N_t(x) \gamma(x, \eta^N_t) r(y, \eta^N_t)
}{
    \varphi^N_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated 
as a randomly chosen individual, we define
a continuous-time jump process 
whose transition rates, conditional on $(\varphi^N_t)_{t=0}^T$, 
are given by \eqref{eqn:informal_rates}. Because we are tracing the
lineage backwards 
in time we make the substitution $s=T-t$ and write
$(L^N_s)_{s=0}^T$ for the location of a lineage that moves 
according to these jump rates.
Then, abusing notation to write 
$q_\theta(x, y)$ for the density of $q_\theta(x,dy)$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^N_{s+ds}) - f(y) \;|\; L^N_s = y]
    \\&\qquad 
    =
    ds\, \theta \int \left(f(x) - f(y)\right)
    \frac{
        \varphi^N_{T-s}(x) \gamma(x, \eta^N_{T-s}) r(y, \eta^N_{T-s})
    }{
        \varphi^N_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
(Note that this integral is with respect to $x$.)
Referring back to Remark~\ref{rem:DG_limit},
a quick calculation shows that as $N \to \infty$,
\begin{align*}
    &
    \theta \int \big(f(x) - f(y)\big) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \big\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \big\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

@@

\begin{proposition}[Line of descent construction]
    \label{prop:limiting_construction}
Define $J(x,\eta)$ and $\beta(x,\eta)$ by
\begin{gather*}
    r(x,\eta)\gamma(x,\eta)\covq(x) = J(x,\eta) J(x,\eta)^T \\
    \beta(x, \eta) = r(x,\eta)\gamma(x,\eta)\big(\meanq(x) + 2 \covq(x) \grad \log r(x, \eta)\big) .
\end{gather*}
Associate with each label $a \in \labelspace = \cup_{k\geq 1}\IN^k$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
and with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process
on $\IR^d \times [0, \infty)$ 
with mean measure $\eta_0\times \lambda$
(the product of $\eta_0$ and Lebesgue measure).
For each $i$, begin a line of descent
with label $i$, location $X_i(0) = x_i$, 
level $U_i(0) = u_i$, and birth time $\tau_i = 0$.

@@

Proofs of the remainder of these results are in Section~\ref{sec:lookdown_proofs}.

@@

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right) .
\end{equation*}
and 
we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function
and $\check{\zeta}(x) = \zeta(-x)$.
The example we have in mind is $\zeta$ (and therefore $\rho$) being the density of
a mean zero Gaussian random variable.  
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}
